\bf Estimating the number of clusters using cross-validation

### 2018-03-01 cvclust.tex
address third AE point

  \item \label{step:transform}
    Compute the eigendecomposition $\hat{\Sigma} = \Gamma\Lambda\Gamma^\T$.
    Choose a random (Haar distributed) $P \times P$ orthogonal matrix $Q$.
    Rescale and rotate the original data matrix $\dataX$ to get a
    transformed data matrix defined by
    \[
      \tilde{\dataX} = \dataX\Gamma\Lambda^{-1/2}Q.
    \]

@@

>>>>

  \item \label{step:transform}
    Compute a decomposition $\hat{\Sigma} = \Gamma\Lambda\Gamma^\T$
    where $\Gamma$ has orthonormal columns and
    $\Lambda$ is diagonal and positive-definite.
    Choose a random (Haar distributed) $P \times P$ orthogonal matrix $Q$.
    Rescale and rotate the original data matrix $\dataX$ to get a
    transformed data matrix defined by
    \[
      \tilde{\dataX} = \dataX\Gamma\Lambda^{-1/2}Q.
    \]

@@

In many applications, including our application in the sequel to a yeast cell
cycle dataset, the number of observations $N$ is below the number of
features~$P$. When this happens, the sample covariance matrix $\hat \Sigma$
will not have full rank~$P$, and will be a poor estimate of the population
covariance matrix. In these situations, though, it will still be the case that
a decomposition $\hat \Sigma = \Gamma \Lambda \Gamma^\T$ exists with $\Gamma$
having orthonormal columns and $\Lambda$ being diagonal and positive-definite.
It will also be the case that $\Lambda^{-1/2}$ and hence $\tilde{\dataX}$
exist. What fails in these situation is that, since $\hat \Sigma$ is no longer
a good estimate of $\Sigma$, it will no longer be the case that our correction
reliably transforms the noise covariance to $I$. Still, in these situations,
applying the correlation correction may give better results than using the
original untransformed data. In practice we inspect the resulting estimate
$\hat k$ both with and without the correlation correction, and if the two
answers agree, we can be more confident in our answer.
